Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-20 paper/paper.tex
integrated more papers

\maketitle
\small{
    \textbf{1} Universit\'e Paris-Saclay, INRIA, Parietal Team, Saclay, France\\
    \textbf{2} Universit\'e Paris-Saclay, CEA, Neurospin, Gif-sur-Yvette, France\\
    \textbf{3} Department of Psychiatry, Psychotherapy and Psychosomatics, RWTH Aachen University, Aachen, Germany\\
    \textbf{4} Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
    \textbf{5} CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
    \textbf{6} University Paris Diderot, Sorbonne Paris Cit\'e, Paris, France\\
    \textbf{7} Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
}

@@

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
  \caption{\textbf{Default mode network: key functions.}
  Neurobiological overview of the DMN with its major constituent nodes and
  their associated functional roles relevant in our functional account.}
  \label{fig:neurobiological_overview_DMN}
\end{figure}
%
\subsection{The posteromedial cortex: global monitoring and information integration}
The midline structures of the DMN,
including the posteromedial cortex (PMC) and
the medial prefrontal cortex (mPFC),
are probably responsible for the highest turn-over
of glucose energy consumption \citep{raichle2001pnas}.
These metabolic characteristics go hand-in-hand with
functional neuroimaging analyses \citep{andrews2010} in
suggesting that the PMC and mPFC
potentially represent the functional backbone of the DMN.

@@

\subsection{The hippocampus: memory, space, and experience replay}
The DMN midline is probably closely linked
with the HC in the medial temporal lobe \citep{vincet2006} --
a region long known to be involved in
memory operations and spatial navigation in animals and humans.

While the HC
is traditionally believed to allow remembering the past,
there is now increasing evidence for a role
in constructing mental models in general
\citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017}.
Its recursive anatomical architecture
may be specifically designed to allow reconstructing
entire episodes of experience from memory fragments.
Indeed,
hippocampal damage is
not only associated with an impairment in reexperiencing the past (i.e., amnesia),
but also imagining one's own future and
fictitious experiences more broadly \citep{hassabis2007patients}.
Mental scenes created by HC-lesioned patients exposed a lack of
spatial integrity, richness in detail, and overall coherence.
%
Single-cell recordings in the animal HC revealed
some constantly active neuronal ensembles whose firing coincided with
specific locations in space during environmental navigation.
London taxi drivers, individuals with high performance in spatial navigation,
were shown to exhibit increased gray-matter volume in the
HC \citep{maguire2000navigation}.
Indeed, when an animal is choosing between alternative
paths, the corresponding neuronal populations in the HC
spike one after another  \citep{johnson2007neural}.
These neuronal patterns appear to also indicate upcoming behavior
\citep{pfeiffer2013hippocampal}.

HC function probably extend beyond simple forms of
encoding and reconstruction of memory and spatial information.
Based on large-scale recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events \citep{buzsaki2004large}.
Specific spiking sequences, which were elicited in experimental task conditions,
have been shown to be reenacted spontaneously during
quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
Moreover, spike sequences measured in hippocampal place cells of rats
featured re-occurrence directly after experimental trials
as well as directly before upcoming experimental trials \citep{diba2007forward}.
Such hippocampal ensembles burt during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage, potentially also in the nodes of the DMN.
These HC-subserved mechanisms
probably contribute to advanced cognitive processes that require
recollecting autobiographical memory episodes, as well as
reexperiencing or newly generating mental scenarios
\citep{hassabis2007patients}.

The HC thus orchestrates experience of environmental aspects for
consolidations based on reenactment and for integration into
rich mental scene construction \citep{deuker2016event, bird2010establishing}.
In this way, the HC may contribute to improving
ongoing perception of and action on the environment
\citep{maguire2016}.
% (TODO: relate to self-play in deepmind Go).
% (TODO: Kai had an interesting idea: the deja-vu effect
% could be when you already accidentally perturbed past
% such that it really occures in reality later.)

@@

\begin{wrapfigure}{r}{.4\linewidth}
  \caption{Reinforcement Learning in a nutshell}  
  \centering
  \includegraphics[width=.9\linewidth]{rl.png}
  \label{fig:rl}
\end{wrapfigure}

@@

\subsubsection{Accumulated rewards and policies}
The behavior of the agent is governed by some kind of \textit{policy}.
It maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory of action cascades / trajectories as follows:
\begin{eqnarray*}
  \begin{split}
    \textbf{choose action: }a_0 &\sim \pi(a|s_0)\\
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1)\\
    \vdots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t})\\
    \vdots
  \end{split}
\end{eqnarray*}

@@

Since an action taken in the present moment might have repercussions in the far future, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account expected reward in the future.
A common approach to
modeling this accumulation is the time-discounted \textit{cumulative reward}% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the DMN ?
    % brain ?
    }
  The amygdala is known to be involved in evaluation of biological significance,
  whereas
  the NAc is involved in reward evaluation. Both are known to
  be connected to the vmPFC of the DMN.
\end{mdframed}
  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$, in an average sense to be made precise latter. In
  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who is solely concerned about its immediate rewards.
Such an agent would have no horizon for the future,
which is not compatible with long-term goal planning
as potentially implemented in the DMN.
To allow a learning process to arise,
it is necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
% The reward discount factor
$\gamma$ can be seen as calibrating risk-seeking valency of the intelligent agent,
that is, the decision related to reward-delay trade-off schedules.
Such an agent puts relatively more emphasis on rewards expected in
a short future and pays less attention to
rewards expected at a later time. Precisely, rewards that are not expected to come within $\tau := 1/(1 - \gamma)$ time steps from the present point are disregarded.
This reduces the variance of expected rewards accumulated across action cascades by limiting the length of this cascades.
% given that there is more uncertainty in the farsighted future.
% It is also important to appreciate that a stochastic policy estimation
% is more advantageous in many RL settings.
% For instance, a deterministic strategy in playing rock-paper-scissors can
% be easily exploited and a uniformly random choice is optimal.

\subsection{The knots and bolts of Reinforcement Learning in the DMN}
\textbf{XXX Elvis: ==> Elvis a short preface is needed here, explaining the roadmap of this subsection from a high-level perspective!}
\subsubsection{The objective value of an action given a state}
Given only the limited information available from a POMDP, at a state $s$ the average quality or usefulness of taking an action $a$ under a policy $\pi$ is completely captured by the single number
% We can now formulate the purpose of the DMN
% as finding a policy $\pi$ for the human agent that maximizes the
% expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a],
\end{equation}
called the $Q$-value for the state-action pair $(s,a)$.
In other words, $Q^{\pi}(s,a)$ corresponds to the expected reward
over all possible action trajectories, in which
the agent sets out in the environment in state
$s$, chooses action $a$, and then follows the policy $\pi$ to select future actions.
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
For the brain,
$Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
utility of executing a specific action; it answers the question
"What is the expected utility of taking action $a$ in this situation ?".
$Q^{\pi}(s,a)$ offers a formalization of optimal behavior that
may well reflect processing aspects of the DMN in human agents.

@@

\begin{mdframed}
  \textbf{ Neural correlates of the Bellman equation in the DMN:}
  ...
\end{mdframed}

@@

  \subsubsection{Self-training and the loss function}
There can be no learning without a loss function in sight.
The action $a$ taken in state $s$ according to the policy matrix defined in
\eqref{eq:policy} triggers a reward $r$ collected by the agent,
% (measured on their internal currency of worldly values ?),
after which the environment transitions to a new state $s' \in \mathcal S$.
% \textbf{XXX Please try to smoothen the previous sentence}
At the end of this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced, which represents an exemplar
behavior of the agent and is recorded in replay memory buffer --which we hypothesize to be the
hippocampus--, possibly discarding the oldest entries to make space:
$\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1} + \delta \theta_{k}$ of the parameters for its approximate model of $Q$-value functional. This warrants learning,
and a loss function is in demand.
% \subsubsection{Stochastic gradient in the DMN via experience replay}
The Bellman equation \eqref{eq:bellman} provides a way to produce such a loss function \eqref{eq:oracle} as we now outline.
% A practically successful RL approach
% needs ways to efficiently approximate the state-action value function
% to allow for scaling to large state and action spaces.
Replay consists in sampling (uniform or importance-weighted
\footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
$(s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$.
The agent then tries to approximate
the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the Bellman equation \eqref{eq:bellman}, namely
\begin{equation}
  y_k := y_k(s,a,s') =  r + \gamma \max_{a'} \tilde{Q}(s', a'|\theta_{k-1}),
\end{equation}
with the prediction of a parametrized regression model $(s,a)
\mapsto \tilde{Q}(s, a|\theta_{k-1})$.
From a neurobiological perspective,
experience replay would be manifested as the re-occurrence of a
neuron spiking sequence that also occurs during specific actions.
%, but the replay has a much faster time scale.
In the human brain, the hippocampus is likely to contribute to such a mechanism,
as neuroscience experiments on various animals have repeatedly indicated
in rats, mice, cats, rabbits, songbirds, and
monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}.

@@

In the human brain,
the HC could contribute to synthesizing imagined sequences of world states,
observations and rewards \citep{aronov2017, chao2017interaction}.
These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics. This would be a simple control algorithm by evaluating all legal actions and selecting the action with
highest expected cumulative rewards.
In POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
\textbf{XXX Elvis: refs needed.}.
Much recent work points to Monte Carlo or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines. \textbf{XXX Elvis: refs needed in last sentence. Plus, should be really bring up Bayesian?} For sampling-based methods it is still almost non-existent beyond the description of receptive fields. \textbf{XXX Elvis: rephrase last sentence}

\subsection{Putting everything together}
The DMN is today known to consistently increase in neural
  activity when humans engage in goal-directed behavior that are detached from
  current sensory environment \citep{kenet2003spontaneously, fiser2004small}
  and it was proposed to be situated at the top of the functional network hierarchies
  \citep{carhart2010default, margulies2016situating}.
  Its putative involvement in thinking about the past,
  the future, and hypothetical possibilities
  appears to tie in with the implicit computation of
  action and state cascades as a function of what happened in the past.
  A policy represents the repertoire of possible actions
    on the world given a current state. It encodes the probabilites of
    choosing actions to be executed in a certain situation.
% Interactions between external and internal sources of
% information are probably commonplace during perception \citep{gershman2010learning}
The DMN may subserve
  constant exploration of possible future actions and their
  cumulative reward outcomes. Implicit computation of future choices
  provides an explanation for the
  evolutionary emergence and practical usefulness of mind-wandering
  in humans.
\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
  \caption{\textbf{DMN: neurobiological implementation of Reinforcement Learning.}
  Overview of how the constituent nodes of the Default Mode Network (DMN)
  map onto computational components necessary for an RL agent.}
  \label{fig:rl_process_chart}
\end{figure}

@@

More specifically,
environmental stimuli especially important for humans are frequently of
social nature. This is probably unsurprising
given that
the complexity of the social systems
is a likely human-defining property
\citep{tomasello2009cultural}.
According to the ``social brain hypothesis'',
the human brain has especially been shaped for
forming and maintaining increasingly complex
social systems,
which allows solving ecological problems by social relationships
\citep{whiten1988machiavellian}.
Indeed, social topics amount to roughly
two thirds of human everyday communication \citep{dunbar1997human},
while
mind-wandering at daytime and dreams during sleep
are rich in stories about people and
the complex relationships between them.
%
In line with this, the DMN has been proposed to reflect
continuous processing of social information as a
physiological baseline of human brain function
\citep{schilbach2008minds}. This view was later challenged by observing
analogues of the DMN in monkeys \citep{mantini2011default}
and rats \citep{lu2012rat}, both species with
supposedly less advanced social-cognitive capacities.
Rather than functional specificity for processing social information,
the present formal account can parsimoniously incorporate
the dominance of social content in
human mental activity
as high value function estimates for information about humans
\citep{baker2009action, kampe2001psychology, krienen2010clan}.
The DMN may thus exert control on the human agent in a way that prioritizes
appraisal of and action towards social contexts,
without entirely neglecting environmental cues of the physical world.
In sum,
our account on the DMN directly implies
its previously proposed ``sentinel'' function
of monitoring the environment for salient information
in general and
inherently incorporates the importance of social environmental cues
as an important special case.

@@

% \paragraph{Comparison to our proposed theory.}
On the surface, a common point between the FEP and our proposed RL-based framework
  is placing the minimization of a surprise signal at the core of brain function.
  Indeed in RL, surprise minimization is subsumed  by accurate prediction of
  rewarding outcomes in the future. A ``free-energy'' agent is barely a biomechanical machine which has the tendency to resist undesired / harmful phase-transitions.
  \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}. Such a theory cannot by itself,
  explain the emergence of strategic behavior inherent in humans (e.g.,  cite dark-room experiment).

@@

\paragraph{Acknowledgment.}
Starwars!
% {\small The research leading to these results has received funding from the
% European Union Seventh Framework Programme (FP7/2007-2013)
% under grant agreement no. 604102 (Human Brain Project).
% Further support was received from
% the German National Academic Foundation (D.B.).
% }

@@

\appendix

>>>>

\maketitle
\small{
    \textbf{1} Universit\'e Paris-Saclay, INRIA, Parietal team, Saclay, France\\
    \textbf{2} Universit\'e Paris-Saclay, CEA, Neurospin, Gif-sur-Yvette, France\\
    \textbf{3} Department of Psychiatry, Psychotherapy and Psychosomatics, RWTH Aachen University, Aachen, Germany\\
    \textbf{4} Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
    \textbf{5} CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
    \textbf{6} University Paris Diderot, Sorbonne Paris Cit\'e, Paris, France\\
    \textbf{7} Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
}

@@

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
  \caption{\textbf{Default mode network: key functions.}
  Neurobiological overview of the DMN with its major constituent nodes and
  their associated functional roles relevant in our functional account.}
  \label{fig:neurobiological_overview_DMN}
\end{figure}
%
\subsection{The posteromedial cortex: global monitoring and information integration}
The midline structures of the DMN,
including the posteromedial cortex (PMC) and
the medial prefrontal cortex (mPFC),
are probably responsible for the highest turn-over
of glucose energy consumption \citep{raichle2001pnas}.
These metabolic characteristics go hand-in-hand with
functional and structural neuroimaging analyses
suggesting that the PMC and mPFC
potentially represent the functional backbone of the DMN
\citep{andrews2010, hagmann2008mapping}.

@@

\subsection{The hippocampus: memory, space, and experience replay}
The DMN midline is probably closely linked
with the HC in the medial temporal lobe \citep{vincet2006, shannon2013morning} --
a region long known to be involved in
memory operations and spatial navigation in animals and humans.

While the HC
is traditionally believed to allow remembering the past,
there is now increasing evidence for a role
in constructing mental models in general
\citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017}.
Its recursive anatomical architecture
may be specifically designed to allow reconstructing
entire episodes of experience from memory fragments.
Indeed,
hippocampal damage is
not only associated with an impairment in reexperiencing the past (i.e., amnesia),
but also imagining one's own future and
imagined experiences more broadly \citep{hassabis2007patients}.
Mental scenes created by HC-lesioned patients exposed a lack of
spatial integrity, richness in detail, and overall coherence.
%
Single-cell recordings in the animal HC revealed
some constantly active neuronal ensembles whose firing coincided with
specific locations in space during environmental navigation.
London taxi drivers, individuals with high performance in spatial navigation,
were shown to exhibit increased gray-matter volume in the
HC \citep{maguire2000navigation}.
Indeed, when an animal is choosing between alternative
paths, the corresponding neuronal populations in the HC
spike one after another  \citep{johnson2007neural}.
These neuronal patterns appear to also indicate upcoming behavior
\citep{pfeiffer2013hippocampal}.

HC function probably extend beyond simple forms of
encoding and reconstruction of memory and spatial information.
Based on large-scale recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events \citep{buzsaki2004large}.
Specific spiking sequences, which were elicited in experimental task conditions,
have been shown to be reenacted spontaneously during
quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
Moreover, spike sequences measured in hippocampal place cells of rats
featured re-occurrence directly after experimental trials
as well as directly before upcoming experimental trials \citep{diba2007forward}.
Such hippocampal ensembles burt during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage, potentially also in the nodes of the DMN.
Moreover, invasively triggering experience recall
in the HC of mice during sleep
has been shown to subsequently alter action choice during wakefulness
\citep{lavilleon2015}.
These HC-subserved mechanisms
probably contribute to advanced cognitive processes that require
recollecting autobiographical memory episodes, as well as
reexperiencing or newly generating mental scenarios
\citep{hassabis2007patients}.

The HC thus orchestrates experience of environmental aspects for
consolidations based on reenactment and for integration into
rich mental scene construction \citep{deuker2016event, bird2010establishing}.
In this way, the HC may contribute to improving
ongoing perception of and action on the environment
\citep{maguire2016, lavilleon2015}.
% (TODO: relate to self-play in deepmind Go).
% (TODO: Kai had an interesting idea: the deja-vu effect
% could be when you already accidentally perturbed past
% such that it really occures in reality later.)

@@

\begin{wrapfigure}{r}{.4\linewidth}
  \centering
  \includegraphics{rl.png}
  \caption{RL in a nutshell}
  \label{fig:rl}
\end{wrapfigure}

@@

\subsubsection{Accumulated rewards and policies}
The behavior of the agent is governed by some kind of \textit{policy}.
It maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory of action cascades / trajectories as follows
\begin{eqnarray*}
  \begin{split}
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1)\\
    \ldots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}

@@

Since an action taken in the present moment $t$
might have repercussions in the far future, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account expected reward in the future.
A common approach to
modeling this accumulation is the time-discounted \textit{cumulative reward}% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the DMN ?
    % brain ?
    }
  The amygdala is known to be involved in evaluation of biological significance,
  whereas
  the NAc is involved in reward evaluation. Both are known to
  be connected to the vmPFC of the DMN.
\end{mdframed}
  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$, in an average sense to be made precise latter. In
  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who is solely concerned about its immediate rewards.
Such an agent would have no horizon for the future,
which is not compatible with long-term goal planning
as potentially implemented in the DMN.
To allow a learning process to arise,
it is necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
% The reward discount factor
$\gamma$ can be seen as calibrating risk-seeking valency of the intelligent agent,
that is, the decision related to reward-delay trade-off schedules.
Such an agent puts relatively more emphasis on rewards expected in
a short future and pays less attention to
rewards expected at a later time. Precisely, rewards that are not expected to come within $\tau := 1/(1 - \gamma)$ time steps from the present point are disregarded.
This reduces the variance of expected rewards accumulated across action cascades by limiting the length of this cascades.
% given that there is more uncertainty in the farsighted future.
% It is also important to appreciate that a stochastic policy estimation
% is more advantageous in many RL settings.
% For instance, a deterministic strategy in playing rock-paper-scissors can
% be easily exploited and a uniformly random choice is optimal.

\subsection{The knots and bolts for RL in the DMN}
\subsubsection{The objective value of an action given a state}
Given only the limited information available from a POMDP, at a state $s$ the average quality or usefulness of taking an action $a$ under a policy $\pi$ is completely captured by the single number
% We can now formulate the purpose of the DMN
% as finding a policy $\pi$ for the human agent that maximizes the
% expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a],
\end{equation}
called the $Q$-value for the state-action pair $(s,a)$.
In other words, $Q^{\pi}(s,a)$ corresponds to the expected reward
over all possible action trajectories, in which
the agent sets out in the environment in state
$s$, chooses action $a$, and then follows the policy $\pi$ to select future actions.
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
For the brain,
$Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
utility of executing a specific action; it answers the question
"What is the expected utility of taking action $a$ in this situation ?".
$Q^{\pi}(s,a)$ offers a formalization of optimal behavior that
may well reflect processing aspects of the DMN in human agents.

@@

\begin{mdframed}
  \textbf{ NCs for the Bellman equation:}
  ...
\end{mdframed}

@@

\subsubsection{The loss function}
There can be no learning without a loss function in sight.
The action $a$ taken in state $s$ according to the policy matrix defined in
\eqref{eq:policy} triggers a reward $r$ collected by the agent,
% (measured on their internal currency of worldly values ?),
after which the environment transitions to a new state $s' \in \mathcal S$.
% \textbf{XXX Please try to smoothen the previous sentence}
At the end of this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced, which represents an exemplar
behavior of the agent and is recorded in replay memory buffer --which we hypothesize to be the
hippocampus--, possibly discarding the oldest entries to make space:
$\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1} + \delta \theta_{k}$ of the parameters for its approximate model of $Q$-value functional. This warrants learning,
and a loss function is in demand.
% \subsubsection{Stochastic gradient in the DMN via experience replay}
The Bellman equation \eqref{eq:bellman} provides a way to produce such a loss function \eqref{eq:oracle} as we now outline.
% A practically successful RL approach
% needs ways to efficiently approximate the state-action value function
% to allow for scaling to large state and action spaces.
Replay consists in sampling (uniform or importance-weighted
\footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
$(s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$.
The agent then tries to approximate
the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the Bellman equation \eqref{eq:bellman}, namely
\begin{equation}
  y_k := y_k(s,a,s') =  r + \gamma \max_{a'} \tilde{Q}(s', a'|\theta_{k-1}),
\end{equation}
with the prediction of a parametrized regression model $(s,a)
\mapsto \tilde{Q}(s, a|\theta_{k-1})$.
From a neurobiological perspective,
experience replay would be manifested as the re-occurrence of a
neuron spiking sequence that also occurs during specific actions.
%, but the replay has a much faster time scale.
In the human brain, the hippocampus is likely to contribute to such a mechanism,
as neuroscience experiments on various animals have repeatedly indicated
in rats, mice, cats, rabbits, songbirds, and
monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}.

@@

In the human brain,
the HC could contribute to synthesizing imagined sequences of world states,
observations and rewards \citep{aronov2017, chao2017interaction}.
These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics \citep{lavilleon2015}.
This would be a simple control algorithm by evaluating all legal actions and selecting the action with
highest expected cumulative rewards.
In POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
\textbf{XXX Elvis: refs needed.}.
Much recent work points to Monte Carlo or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines. \textbf{XXX Elvis: refs needed in last sentence. Plus, should be really bring up Bayesian?} For sampling-based methods it is still almost non-existent beyond the description of receptive fields. \textbf{XXX Elvis: rephrase last sentence}

% \paragraph{Does hippocampal replay equal inverse reinforcement-learning?}
% %\begin{mdframed}
%   Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
%   a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
%   (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
%   \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
%   (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
%   For instance, given traces of motor actions from an adult to grab a cup on a table,
%   can an observing child figure out what ``langragian'' functional is being minimized by the former?
%   How? Can they reproduce this optimal behavior?
% %\end{mdframed}
% Such questions are of course pertinent to our decision-making theory for the DMN.
% In the general artificial case, the problem has been extensively studied and partially solved by
% \citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
% ``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...

% IRL is is suited for problems in which it's hard to define what the reward function could be
% (e.g.,  car-driving, drone maneuvering, etc.) ...

\subsection{Putting everything together}
The DMN is today known to consistently increase in neural
  activity when humans engage in goal-directed behavior that are detached from
  current sensory environment \citep{kenet2003spontaneously, fiser2004small}
  and it was proposed to be situated at the top of the functional network hierarchies
  \citep{carhart2010default, margulies2016situating}.
  Its putative involvement in thinking about the past,
  the future, and hypothetical possibilities
  appears to tie in with the implicit computation of
  action and state cascades as a function of what happened in the past.
  A policy represents the repertoire of possible actions
    on the world given a current state. It encodes the probabilites of
    choosing actions to be executed in a certain situation.
% Interactions between external and internal sources of
% information are probably commonplace during perception \citep{gershman2010learning}
The DMN may subserve
  constant exploration of possible future actions and their
  cumulative reward outcomes. Implicit computation of future choices
  provides an explanation for the
  evolutionary emergence and practical usefulness of mind-wandering
  in humans.
\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
  \caption{\textbf{Default mode network: reinforcement learning implementation.}
  Overview of how the constituent nodes of the DMN
  map onto computational components necessary for a Markov decision process.}
  \label{fig:rl_process_chart}
\end{figure}

@@

More specifically,
environmental stimuli especially important for humans are frequently of
social nature. This is probably unsurprising
given that
the complexity of the social systems
is a likely human-defining property
\citep{tomasello2009cultural}.
According to the ``social brain hypothesis'',
the human brain has especially been shaped for
forming and maintaining increasingly complex
social systems,
which allows solving ecological problems by social relationships
\citep{whiten1988machiavellian}.
Indeed, social topics amount to roughly
two thirds of human everyday communication \citep{dunbar1997human},
while
mind-wandering at daytime and dreams during sleep
are rich in stories about people and
the complex relationships between them.
%
In line with this, the DMN has been proposed to reflect
continuous processing of social information as a
physiological baseline of human brain function
\citep{schilbach2008minds}. This view was later challenged by observing
analogues of the DMN in monkeys \citep{mantini2011default},
cats \citep{popa2009contrasting},
and rats \citep{lu2012rat}, three species with
supposedly less advanced social-cognitive capacities.
Rather than functional specificity for processing social information,
the present formal account can parsimoniously incorporate
the dominance of social content in
human mental activity
as high value function estimates for information about humans
\citep{baker2009action, kampe2001psychology, krienen2010clan}.
The DMN may thus exert control on the human agent in a way that prioritizes
appraisal of and action towards social contexts,
without entirely neglecting environmental cues of the physical world.
In sum,
our account on the DMN directly implies
its previously proposed ``sentinel'' function
of monitoring the environment for salient information
in general and
inherently incorporates the importance of social environmental cues
as an important special case.

@@

\subsubsection{Comparison to our proposed theory}
On the surface, a common point between the FEP and our proposed RL-based framework
  is placing the minimization of a surprise signal at the core of brain function.
  Indeed in RL, surprise minimization is subsumed  by accurate prediction of
  rewarding outcomes in the future. A ``free-energy'' agent is barely a biomechanical machine which has the tendency to resist undesired / harmful phase-transitions.
  \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}. Such a theory cannot by itself,
  explain the emergence of strategic behavior inherent in humans (e.g.,  cite dark-room experiment).

@@

\paragraph{Acknowledgment}
% {\small The research leading to these results has received funding from the
% European Union Seventh Framework Programme (FP7/2007-2013)
% under grant agreement no. 604102 (Human Brain Project).
% Further support was received from
% the German National Academic Foundation (D.B.).
% }

@@

\appendix
\section{Free-energy principles!}
The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making. This ideas were first developed by Hinton et al. in the early 90s in building their Helmholtz machine. Theories like Friston's free-energy principle and active-inference will then emerge as particular instances of this general framework, with particular design choises. For instance, the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
